Synthesising Termination Proofs for Bit-Vector Programs

### 2014-07-06 papers/termination/paper.tex
nudge related

\section{Conclusions and Related Work}
%\subsection{Termination analysis}

Automated program termination is a research topic that has received a fair
amount of attention from the software verification community.  In order to
compare our technique to the rest of the area, Figure~\ref{fig:handletable}
summarises the related works with respect to the assumptions they make about
programs and ranking functions.

Somehow expected, most of the techniques are specialised in the synthesis of
linear ranking functions for linear programs over integers (or rationals)
\cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/LeeWY12,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,
DBLP:conf/atva/HeizmannHLP13,DBLP:conf/cav/BradleyMS05,DBLP:conf/tacas/CookSZ13,DBLP:conf/cav/KroeningSTW10}. 
Among them, Lee et al.~make use of transition predicate abstraction,
algorithmic learning, and decision procedures
%to compute transition invariants as proofs for the termination of linear programs 
\cite{DBLP:conf/cav/LeeWY12},
Leike and Heizmann propose linear ranking templates 
%in order to synthesise
%termination arguments for linear loop programs 
\cite{DBLP:conf/tacas/LeikeH14}, whereas
Bradley et al.~compute lexicographic linear ranking functions supported by inductive linear invariants 
%for loops with linear guards and transitions: 
\cite{DBLP:conf/cav/BradleyMS05}. 
%Learning: \cite{DBLP:conf/cav/HeizmannHP14}

While the synthesis of termination arguments for linear programs over
integers is indeed well-covered in the literature, there is very limited
work for programs over machine integers.  Cook et al.~present a method
based on a reduction to Presburger arithmetic, and a template-matching
approach for predefined classes of ranking functions based on reduction to
SAT- and QBF-solving~\cite{DBLP:conf/tacas/CookKRW10}.  Similarly, the only
work we are aware of that can compute non-linear ranking functions for
imperative loops with polynomial guards and polynomial assignments
is~\cite{DBLP:conf/vmcai/BradleyMS05}.  However, this work extends only to
polynomials.

Given the lack of research in handling not-linear programs, as well
as programs over bit vectors and floats, our work focused on covering
these areas.  One of the obvious conclusions that can be reached from
observing Figure~\ref{fig:handletable}, is that most of the works tend to
specialise on a certain aspect of termination proving that they can solve
efficiently.  Conversely to this view, we aim for generality, as we
do not restrict the form of the synthesised ranking functions, nor the form
of the input programs.

As mentioned in Section~\ref{sec:intro}, approaches based on Ramsey's
theorem compute a set of local termination conditions that decrease as
execution follows through the loop and require expensive reachability
analyses~\cite{DBLP:conf/lpe/CodishG03,DBLP:conf/lics/PodelskiR04,DBLP:conf/pldi/CookPR06}. 
In an attempt to reduce the complexity of checking the validity of the
termination argument, Cook et al.~present an iterative termination proving
procedure that searches for lexicographic termination
arguments~\cite{DBLP:conf/tacas/CookSZ13}, whereas Kroening et
al.~strengthen the termination argument such that it becomes a transitive
relation~\cite{DBLP:conf/cav/KroeningSTW10}.

Proving program termination implies the simultaneous search for a
termination argument and a supporting invariant.  
%For this purpose, several tools alternate between calls to a safety prover and calls to a ranking function synthesis tool \cite{}.  
%In order to share more information about
%the state of the termination proofs between the two tools, 
Brock\-schmidt et al.~share the same representation of the state of the termination proof
between the safety prover and the ranking function synthesis tool \cite{DBLP:conf/cav/BrockschmidtCF13}.  
In \cite{DBLP:conf/cav/BradleyMS05},
Bradley et al.~combine the generation of ranking functions with the
generation of invariants to form one constraint solving problem such that
the necessary supporting invariants for the ranking function are discovered
on demand.  In our setting, both the ranking function and the supporting
invariant are iteratively constructed through the same refinement loop.

>>>>

\section{Conclusions and Related Work}

There has been substantial prior work on automated program termination
analysis.  Figure~\ref{fig:handletable} summarises the related work with
respect to the assumptions they make about programs and ranking functions. 
Most of the techniques are specialised in the synthesis of linear ranking
functions for linear programs over integers (or
rationals)~\cite{DBLP:conf/pldi/CookPR06, DBLP:conf/cav/LeeWY12,
DBLP:conf/popl/Ben-AmramG13, DBLP:conf/vmcai/P04,
DBLP:conf/atva/HeizmannHLP13, DBLP:conf/cav/BradleyMS05,
DBLP:conf/tacas/CookSZ13, DBLP:conf/cav/KroeningSTW10}.  Among them, Lee et
al.~make use of transition predicate abstraction, algorithmic learning, and
decision procedures~\cite{DBLP:conf/cav/LeeWY12}, Leike and Heizmann propose
linear ranking templates~\cite{DBLP:conf/tacas/LeikeH14}, whereas Bradley et
al.~compute lexicographic linear ranking functions supported by inductive
linear invariants~\cite{DBLP:conf/cav/BradleyMS05}.

While the synthesis of termination arguments for linear programs over
integers is indeed well covered in the literature, there is very limited
work for programs over machine integers.  Cook et al.~present a method
based on a reduction to Presburger arithmetic, and a template-matching
approach for predefined classes of ranking functions based on reduction to
SAT- and QBF-solving~\cite{DBLP:conf/tacas/CookKRW10}.  Similarly, the only
work we are aware of that can compute non-linear ranking functions for
imperative loops with polynomial guards and polynomial assignments
is~\cite{DBLP:conf/vmcai/BradleyMS05}.  However, this work extends only to
polynomials.

Given the lack of research in handling not-linear programs, as well as
programs over bit vectors and floats, our work focused on covering these
areas.  One of the obvious conclusions that can be reached from
Figure~\ref{fig:handletable} is that most methods tend to specialise on a
certain aspect of termination proving that they can solve efficiently. 
Conversely to this view, we aim for generality, as we do not restrict the
form of the synthesised ranking functions, nor the form of the input
programs.

As mentioned in Section~\ref{sec:intro}, approaches based on Ramsey's
theorem compute a set of local termination conditions that decrease as
execution proceeds through the loop and require expensive reachability
analyses~\cite{DBLP:conf/lpe/CodishG03, DBLP:conf/lics/PodelskiR04,
DBLP:conf/pldi/CookPR06}.  In an attempt to reduce the complexity of
checking the validity of the termination argument, Cook et al.~present an
iterative termination proving procedure that searches for lexicographic
termination arguments~\cite{DBLP:conf/tacas/CookSZ13}, whereas Kroening et
al.~strengthen the termination argument such that it becomes a transitive
relation~\cite{DBLP:conf/cav/KroeningSTW10}.

Proving program termination implies the simultaneous search for a
termination argument and a supporting invariant.  Brock\-schmidt et
al.~share the same representation of the state of the termination proof
between the safety prover and the ranking function synthesis
tool~\cite{DBLP:conf/cav/BrockschmidtCF13}. 
In~\cite{DBLP:conf/cav/BradleyMS05}, Bradley et al.~combine the generation
of ranking functions with the generation of invariants to form a single
constraint solving problem such that the necessary supporting invariants for
the ranking function are discovered on demand.  In our setting, both the
ranking function and the supporting invariant are iteratively constructed in
the same refinement loop.
